---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### This tutorial will show you how to use the pso() function from particleSwarm, and show time and output comparisons to the psoptim() function from the pso library.


```{r setup}
library(particleSwarm)
```
 
### How to use pso() 

First, we will create a sample function which we will aim to minimize. Note that the input x contains all 3 values for d, D, and N.
```{r}
fitness <- function(x){
  d <- x[1]
  D <- x[2]
  N <- round(x[3])
  
  # define fitness function
  fitness_value <- (N+2)*D*d^2
  
  #define constraint
  g1 <- 1 - D^3*N/(71785*d^4)
  g2 <- (4*D^2-d*D)/(12566*(D * d^3 - d^4)) + 1/(5108 * d^2) - 1
  g3 <- 1 - (140.45*d)/(D^2*N)
  g4 <- (D+d)/1.5 - 1
  
  #penalized constraint violation
  fitness_value <- ifelse( g1 <= 0 & g2 <= 0 & g3 <= 0 & g4 <= 0, fitness_value, fitness_value + 1e3 )
  
  return(fitness_value)
}

# Source: https://rpubs.com/Argaadya/intro-PSO
```

Then, we input our fitness function to pso().
```{r}
pso(
  fn = fitness,
  S = 100,
  lb = c(0.05, 0.25, 2),
  ub = c(2, 1.3, 15),
  w = .5,
  max_iter = 100
)
```

As we can tell from the output parameters, there is an issue with the function pso() because it is returning parameter values outside of the given ranges for the inputs. However, the rest of the function seems to be working correctly. We see that the minimum function value found based on the (incorrect) inputs is very small, and the number of iterations matches our max_iter.

### Performance comparison

Here is an example of the same fitness function and same parameters being minimized with the particle swarm optimizer psoptim() from the pso library.
```{r}
psoptim(rep(NA,3), fn = fitness, lower = c(0.05, 0.25, 2), upper = c(2, 1.3, 15), control = list(maxit = 100, s = 100, w = 0.5, c.p = 1, c.g = 1))
```

This algorithm actually works correctly, so clearly it performs better in terms of accuracy. 

Now, let's run a time comparison.
```{r}
bench::mark(psoptim(rep(NA,3), fn = fitness, lower = c(0.05, 0.25, 2), upper = c(2, 1.3, 15), control = list(maxit = 100, s = 100, w = 0.5, c.p = 1, c.g = 1)))

bench::mark(pso(
  fn = fitness,
  S = 100,
  lb = c(0.05, 0.25, 2),
  ub = c(2, 1.3, 15),
  w = .5,
  max_iter = 100
))
```

We see that the pso() function runs faster, but uses more memory.
